Tutorial_3a

### 2024-12-09 Tutorial_2a.Rmd
update 2a
update tuts

Prepare the data file, and specify priors. The skew_normal prior is set for each grand intercept on the cloglog scale, and reflects our prior belief that hazard values above .5 are less likely than values below .5.

@@

@@

@@

@@

@@

Following Heiss (2021), we can plot the expected value of the posterior predictive distribution -- the predicted hazard values -- for the average participant, for each participant in the data set, and for a brand new hypothetical participant.

@@

```{r epred-grand-plot}
 ggplot(epreds_grand, aes(x=timebin, y=.epred, 
                          fill=prime, color=prime)) +
    stat_lineribbon(point_interval="median_qi",.width=.95,alpha = 0.5) +
    scale_fill_brewer(palette = "Dark2") +
    scale_color_brewer(palette = "Dark2") +
    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
    theme(axis.text.x = element_text(angle=90)) +
    scale_y_continuous(limits=c(0,.6)) +
    labs(y = "predicted hazard") +
    ggtitle("Average participant")
```

```{r save-epred-grand, eval=F}
ggsave("Tutorial_2_Bayesian/figures/M1i_pred_grand.png", width = 10, height = 8, dpi = 800)
```

@@

```{r epred-pid-plot}
 ggplot(epreds_pid, aes(x=timebin, y=.epred, 
                       fill=prime, color=prime)) +
    stat_lineribbon(point_interval="median_qi",
                    .width=.95,
                    alpha = 0.5) +
    scale_fill_brewer(palette = "Dark2") +
    scale_color_brewer(palette = "Dark2") +
    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
    theme(axis.text.x = element_text(angle=90)) +
    scale_y_continuous(limits=c(0,1)) +
    labs(y = "predicted hazard") +
    ggtitle("Each participant (N = 6)") +
    facet_wrap(~pid, ncol=3)
```

```{r save-epred-pid, eval=F}
ggsave("Tutorial_2_Bayesian/figures/M1i_pred_pid.png", width = 10, height = 8, dpi = 800)
```

Third, for a completely new hypothetical participant.

```{r epred-new-pid}
epreds_pid_new <- dat_M1i %>% 
  data_grid(timebin, prime) %>%
  mutate(pid = 7) %>%
  add_epred_draws(model_M1i, 
                  re_formula = NULL, 
                  allow_new_levels=T,    
                  sample_new_levels = "gaussian") %>% 
  mutate(prime = factor(prime, 
                        levels=c(1,2,3),
                        labels=c("blank","congruent","incongruent"))) %>%
  ungroup()
```

```{r epred-new-pid-summary}
epreds_pid_new %>% 
  group_by(pid,timebin,prime) %>% 
  median_qi(.width=.95)
```

```{r epred-new-pid-plot}
 ggplot(epreds_pid_new, aes(x=timebin, y=.epred, 
                       fill=prime, color=prime)) +
    stat_lineribbon(point_interval="median_qi",
                    .width=.95,
                    alpha = 0.5) +
    scale_fill_brewer(palette = "Dark2") +
    scale_color_brewer(palette = "Dark2") +
    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
    theme(axis.text.x = element_text(angle=90)) +
    scale_y_continuous(limits=c(0,1)) +
    labs(y = "predicted hazard") +
    ggtitle("A brand new hypothetical participant") 
```

```{r save-epred-pid-new, eval=F}
ggsave("Tutorial_2_Bayesian/figures/M1i_pred_pid_new.png", width = 10, height = 8, dpi = 800)
```

Fourth, six completely new hypothetical participants.

```{r epred-new-pid-6}
epreds_pid_new6 <- dat_M1i %>% 
  data_grid(timebin, prime, pid = 7:12) %>%
  add_epred_draws(model_M1i, 
                  re_formula = NULL, 
                  allow_new_levels=T,    
                  sample_new_levels = "gaussian") %>% 
  mutate(prime = factor(prime, 
                        levels=c(1,2,3),
                        labels=c("blank","congruent","incongruent"))) %>%
  ungroup()
```

```{r epred-new-pid-6-summary}
epreds_pid_new6 %>% 
  group_by(pid,timebin,prime) %>% 
  median_qi(.width=.95)
```

```{r epred-new-pid-6-plot}
 ggplot(epreds_pid_new6, aes(x=timebin, y=.epred, 
                       fill=prime, color=prime)) +
    stat_lineribbon(point_interval="median_qi",
                    .width=.95,
                    alpha = 0.5) +
    scale_fill_brewer(palette = "Dark2") +
    scale_color_brewer(palette = "Dark2") +
    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
    theme(axis.text.x = element_text(angle=90)) +
    scale_y_continuous(limits=c(0,1)) +
    labs(y = "predicted hazard") +
    ggtitle("Six new hypothetical participants") +
    facet_wrap(~pid, ncol=3)
```

```{r save-epred-pid-6-new, eval=F}
ggsave("Tutorial_2_Bayesian/figures/M1i_pred_pid_new_6.png", width = 10, height = 8, dpi = 800)
```

@@

Summarize and plot.

@@

@@

```{r epred-grand-diff-plot}
 ggplot(epreds_grand_diffs, aes(x=timebin, y=diff, 
                       fill=contrast)) +
  stat_lineribbon(alpha = 0.5, 
                  point_interval = "mean_qi",
                  .width=c(.8,.95)) +
  geom_hline(yintercept = 0, 
             color = "red", 
             lty = 3) +
  scale_y_continuous(limits=c(-0.5,0.3)) +
  scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
  scale_fill_brewer(palette = "Dark2") +
  scale_color_brewer(palette = "Dark2") +
  theme(axis.text.x = element_text(angle=90)) +
  labs(y = "difference in predicted hazard") +
  ggtitle("Grand average marginal effect") +
  facet_wrap(~contrast)
```

@@

```{r epred-pid-diff-plot}
 ggplot(epreds_pid_diffs, aes(x=timebin, y=diff, 
                              fill=contrast, color=contrast)) +
    stat_lineribbon(alpha = 0.5, 
                    point_interval = "mean_qi",
                    .width=c(.8,.95)) +
    #stat_halfeye(point_interval = "mean_qi",.width=c(.8,.95)) +
    geom_hline(yintercept = 0, color = "red", lty = 3) +
    scale_fill_brewer(palette = "Dark2") +
    scale_color_brewer(palette = "Dark2") +
    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
    theme(axis.text.x = element_text(angle=90)) +
    scale_y_continuous(limits=c(-0.6,0.6)) +
    labs(y = "difference in predicted hazard") +
    ggtitle("Subject-specific AMEs") +
    facet_wrap(~pid, ncol=3)
```

@@

[[remove next section and 2 code chunks....]]
And check whether an average of subject-specific AMEs is equal to the grand AME... NO !!!

```{r}
epreds_pid_diffs_av <- epreds_pid_diffs %>% group_by(timebin,contrast,.draw) %>%
  mutate(average_condmeans = mean(diff))
```

```{r}
 ggplot(epreds_pid_diffs_av, aes(x=timebin, y=diff, 
                       fill=contrast)) +
  stat_lineribbon(alpha = 0.5, point_interval = "mean_qi",.width=c(.8,.95)) +
 # stat_halfeye(point_interval = "mean_qi",.width=c(.8,.95))+
  geom_hline(yintercept = 0, color = "red", lty = 3) +
  scale_y_continuous(limits=c(-0.5,0.3)) +
  scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
  scale_fill_brewer(palette = "Dark2") +
  scale_color_brewer(palette = "Dark2") +
  theme(axis.text.x = element_text(angle=90)) +
  labs(y = "difference in predicted hazard") +
  ggtitle("Average of subject-specific AMEs") +
  facet_wrap(~contrast)
```

@@

@@

>>>>

Prepare the data file, and specify priors. The skew_normal prior is set for each grand intercept on the cloglog scale, and reflects our prior belief that hazard values above .5 are less likely than values below .5. 
A prior predictive check is performed in section 8.1.

@@

Fit model M0i.

@@

Prepare the data file, and specify priors. To find the priors, we performed a prior predictive check which is reported in section 8.2.

@@

Fit model M0c.

@@

#
```{r plot-post-M1i-alt}
tidy_fixed <- post %>% 
  select(starts_with("b_"), .chain, .iteration, .draw) %>% 
  rename(chain=.chain, iter=.iteration, draw=.draw) %>% 
  pivot_longer(-c(chain, draw, iter)) %>% 
  mutate(timebin = str_sub(name,10,11),
         timebin = factor(str_remove(timebin,":"),levels=c(6:15)),
         condition = str_sub(name,17,18),
         condition = factor(str_remove(condition,"e"),
                            levels=c(1,2,3),
                            labels=c("blank","congruent","incongruent")))
head(tidy_fixed)
tail(tidy_fixed)

# plot
p_tidy_fixed <- ggplot(tidy_fixed, aes(x = timebin, y = value, fill=condition)) +  
  stat_halfeye(point_interval="median_qi", 
               .width = c(0.80,0.95),
               alpha=0.6,
               position=pd) +
  labs(title = 'Posterior distributions for population-level\neffects in Model M1i',
       x = "time bin", y = "cloglog-hazard") +
  scale_fill_brewer(palette = "Dark2") +
  scale_color_brewer(palette = "Dark2") +

  scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
   theme(axis.text.x = element_text(angle=90)) 
#+
#  facet_wrap(~condition)
p_tidy_fixed
```

```{r save-post, eval=F}
ggsave("Tutorial_2_Bayesian/figures/M1i_postdistr_alt3.png", width = 10, height = 8, dpi = 800)
```

#

@@

Following Heiss (2021), we can plot the expected value of the posterior predictive distribution -- the predicted hazard values -- for the average participant, and for each participant in the data set.

@@

```{r epred-grand-plot}
 p1 <- ggplot(epreds_grand, aes(x=timebin, y=.epred, 
                          fill=prime, color=prime)) +
    stat_lineribbon(point_interval="median_qi",.width=.95,alpha = 0.5) +
    scale_fill_brewer(palette = "Dark2") +
    scale_color_brewer(palette = "Dark2") +
    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
    theme(axis.text.x = element_text(angle=90)) +
    scale_y_continuous(limits=c(0,.7)) +
    labs(y = "predicted hazard")

p1
```

@@

```{r epred-pid-plot}
p2 <- ggplot(epreds_pid, aes(x=timebin, y=.epred, 
                       fill=prime, color=prime)) +
    stat_lineribbon(point_interval="median_qi",
                    .width=.95,
                    alpha = 0.5) +
    scale_fill_brewer(palette = "Dark2") +
    scale_color_brewer(palette = "Dark2") +
    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
    theme(axis.text.x = element_text(angle=90)) +
    scale_y_continuous(limits=c(0,1)) +
    labs(y = "predicted hazard") +
   # ggtitle("Each participant (N = 6)") +
    facet_wrap(~pid, ncol=3)

p2
```

Combine the plots.

```{r save-epred-combined, eval=F}
# combine 2 plots
p1_theme <- p1 + theme(legend.position = "none")

(p1_theme / p2) +
    plot_annotation(tag_levels = 'A') +
    plot_layout(guides = "collect",
                axes = "collect_x") & 
    theme(legend.position = "bottom")
```

Save plot.

```{r save-combined-pred, eval=F}
# save combined plot
ggsave("Tutorial_2_Bayesian/figures/M1i_pred_combined.png", width = 10, height = 12, dpi = 800)
```

@@

Summarize.

@@

Save summary as table.

@@

```{r epred-grand-diff-plot}
 p11 <- ggplot(epreds_grand_diffs, aes(x=timebin, y=diff, 
                       fill=contrast)) +
  stat_lineribbon(alpha = 0.4, 
                  point_interval = "mean_qi",
                  .width=c(.8,.95)) +
  geom_hline(yintercept = 0, 
             color = "red", 
             lty = 3) +
  scale_y_continuous(limits=c(-0.5,0.3)) +
  scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
  scale_fill_brewer(palette = "Dark2") +
  scale_color_brewer(palette = "Dark2") +
  theme(axis.text.x = element_text(angle=90)) +
  labs(y = "difference in predicted hazard") +
  #ggtitle("Grand average marginal effect") +
  facet_wrap(~contrast)

p11
```

@@

```{r epred-pid-diff-plot}
 p22 <- ggplot(epreds_pid_diffs, aes(x=timebin, y=diff, 
                              fill=contrast, color=contrast)) +
    stat_lineribbon(alpha = 0.4, 
                    point_interval = "mean_qi",
                    .width=c(.8,.95)) +
    #stat_halfeye(point_interval = "mean_qi",.width=c(.8,.95)) +
    geom_hline(yintercept = 0, color = "red", lty = 3) +
    scale_fill_brewer(palette = "Dark2") +
    scale_color_brewer(palette = "Dark2") +
    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
    theme(axis.text.x = element_text(angle=90)) +
    scale_y_continuous(limits=c(-0.6,0.6)) +
    labs(y = "difference in predicted hazard") +
    #ggtitle("Subject-specific AMEs") +
    facet_wrap(~pid, ncol=3)

p22
```

@@

Combine the plots.

```{r save-ames-combined, eval=F}
# combine 2 plots
p11_theme <- p11 + theme(legend.position = "none")

(p11_theme / p22) +
    plot_annotation(tag_levels = 'A') +
   # plot_layout(guides = "collect",
  #              axes = "collect_x") & 
    theme(legend.position = "bottom")
```

Save plot.

```{r save-combined-ames, eval=F}
# save combined plot
ggsave("Tutorial_2_Bayesian/figures/M1i_ame_combined.png", width = 10, height = 12, dpi = 800)
```

@@

```{r}
kable(table_prop, caption = "Summarizing the posterior distributions of each contrast by their proportion below and above zero. prop_below = proportion below zero; prop_above = proportion above zero.")
```

@@

# 8. Prior predictive checks

## 8.1. Model M0i

## 8.2 Model M0c

### 2024-12-09 Tutorial_2b.Rmd
update tuts

@@

```{r epred-grand-plot}
 ggplot(epreds_grand, aes(x=timebin, y=.epred, 
                          fill=prime, color=prime)) +
    stat_lineribbon(point_interval="median_qi",
                    alpha = 0.5, 
                    .width = c(.8,.95)) +
    scale_fill_brewer(palette = "Dark2") +
    scale_color_brewer(palette = "Dark2") +
    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
    theme(axis.text.x = element_text(angle=90)) +
    scale_y_continuous(limits=c(0,1)) +
    labs(y = "predicted conditional accuracy") +
    ggtitle("Average participant")
```

@@

```{r plot-epred-pid}
 ggplot(epreds_pid, aes(x=timebin, y=.epred, 
                       fill=prime, color=prime)) +
    stat_lineribbon(point_interval="median_qi",
                    alpha = 0.5, 
                    .width = c(.8,.95)) +
    scale_fill_brewer(palette = "Dark2") +
    scale_color_brewer(palette = "Dark2") +
    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
    theme(axis.text.x = element_text(angle=90)) +
    scale_y_continuous(limits=c(0,1)) +
    labs(y = "predicted conditional accuracy") +
    ggtitle("Each participant (N = 6)") +
    facet_wrap(~pid, ncol=3)
```

@@

>>>>

#
```{r plot-post-M0i-ca-alt}
tidy_fixed_ca <- post_ca %>% 
  select(starts_with("b_"), .chain, .iteration, .draw) %>% 
  rename(chain=.chain, iter=.iteration, draw=.draw) %>% 
  pivot_longer(-c(chain, draw, iter)) %>% 
  mutate(timebin = str_sub(name,10,11),
         timebin = factor(str_remove(timebin,":"),levels=c(6:15)),
         condition = str_sub(name,17,18),
         condition = factor(str_remove(condition,"e"),
                            levels=c(1,2,3),
                            labels=c("blank","congruent","incongruent")))
head(tidy_fixed_ca)
tail(tidy_fixed_ca)

# plot
p_tidy_fixed_ca <- ggplot(tidy_fixed_ca, aes(x = timebin, y = value, fill=condition)) +  
  stat_halfeye(point_interval = "median_qi",
               .width = c(.80,.95),
               alpha=0.7,
               position = pd) +
  labs(title = 'Posterior distributions for population-level\neffects in Model M1i_ca',
       x = "time bin", 
       y = "logit-ca") +
  scale_fill_brewer(palette = "Dark2") +
  scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
  scale_y_continuous(limits=c(-10,10)) +
  theme(axis.text.x = element_text(angle=90)) #+
  #facet_wrap(~condition)
p_tidy_fixed_ca
```

```{r save-post-alt, eval=F}
ggsave("Tutorial_2_Bayesian/figures/M1i_ca_postdistr_alt.png", width = 10, height = 8, dpi = 800)
```

#

@@

```{r epred-grand-plot}
 p1 <- ggplot(epreds_grand, aes(x=timebin, y=.epred, 
                          fill=prime, color=prime)) +
    stat_lineribbon(point_interval="median_qi",
                    alpha = 0.5, 
                    .width = c(.8,.95)) +
    scale_fill_brewer(palette = "Dark2") +
    scale_color_brewer(palette = "Dark2") +
    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
    theme(axis.text.x = element_text(angle=90)) +
    scale_y_continuous(limits=c(0,1)) +
    labs(y = "predicted conditional accuracy") #+
    #ggtitle("Average participant")
p1
```

@@

```{r plot-epred-pid}
p2 <- ggplot(epreds_pid, aes(x=timebin, y=.epred, 
                       fill=prime, color=prime)) +
    stat_lineribbon(point_interval="median_qi",
                    alpha = 0.5, 
                    .width = c(.8,.95)) +
    scale_fill_brewer(palette = "Dark2") +
    scale_color_brewer(palette = "Dark2") +
    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
    theme(axis.text.x = element_text(angle=90)) +
    scale_y_continuous(limits=c(0,1)) +
    labs(y = "predicted conditional accuracy") +
   # ggtitle("Each participant (N = 6)") +
    facet_wrap(~pid, ncol=3)
p2
```

@@

Combine the plots.

```{r save-epred-combined, eval=F}
# combine 2 plots
p1_theme <- p1 + theme(legend.position = "none")

(p1_theme / p2) +
    plot_annotation(tag_levels = 'A') +
    plot_layout(guides = "collect",
                axes = "collect_x") & 
    theme(legend.position = "bottom")
```

Save plot.

```{r save-combined-pred, eval=F}
# save combined plot
ggsave("Tutorial_2_Bayesian/figures/M1i_ca_pred_combined.png", width = 10, height = 12, dpi = 800)
```

### 2024-12-09 Tutorial_3a.Rmd
update tuts

```{r extract-estimates}
est <- tibble(brm = fixef(model_M1i)[,c(1,3,4)],
              glmer = tidy(model_M1i_f, effects = c("fixed"), conf.int=T, conf.level = 0.95)[,c(3,7,8)],
              effectname = names(fixef(model_M1i_f)),
              position = 1:30) 
est
```

@@

>>>>

```{r extract-estimates}
est <- tibble(brm = fixef(model_M1i)[,c(1,3,4)],
              glmer = tidy(model_M1i_f, effects = c("fixed"), conf.int=T, conf.level = 0.95)[,c(3,7,8)],
              effectname = names(fixef(model_M1i_f)),
              position = rep(6:15,3),
              prime = factor(rep(1:3, each=10),labels=c("blank","congruent","incongruent")) )
est %>% print(n=30)
```

@@

#
```{r plot-estimates}
# check hex codes for color paletter
mypal <- brewer.pal(8,"Dark2")

# plot
p1 <- ggplot(data = est, aes(x = position)) +
 geom_point(aes( y = brm[,"Estimate"],  
                 color = "c1", 
                 shape = "s1"),
                 size = 1.5) +
 geom_point(aes( y = glmer$estimate, 
                 color = "c2", 
                 shape = "s2"),
                 size = 1.5, 
                 position = position_nudge(x = .2))  +
 geom_linerange(aes(ymin=brm[,"Q2.5"], ymax=brm[,"Q97.5"], 
                    color = "c1"), 
                show.legend = F) +
 geom_linerange(aes(ymin=glmer$conf.low, ymax=glmer$conf.high, 
                    color = "c2"), 
                show.legend = F, 
                position = position_nudge(x = .2)) +
# geom_hline(yintercept=0,colour = "grey", linetype = 2) +
 scale_x_continuous(breaks = 6:15, labels = 6:15*40) +
 scale_y_continuous(limits=c(-9,3), breaks = c(-9,-6,-3,0,3)) +

 labs( x = "bin endpoint", y = "cloglog-hazard" ) +
 scale_color_manual(name = "Model", 
                                breaks = c("c1", "c2"), 
                                values = cols,
                                labels = c("brm", "glmer")) +
 scale_shape_manual(name = "Model", 
                              breaks = c("s1", "s2"),
                              values = shapes,
                              labels = c("brm", "glmer")) +
  theme(panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle=90)) +
  facet_wrap(~prime)
p1
```

```{r save-plot-comparison, eval=F}
ggsave("Tutorial_3_Frequentist/Comparison.png", height=8, width=10,dpi=800)
```
